Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/compilation.tex
editing pass over compilation section.

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}
  \hdr{\large Topology}{}
  \vspace*{-2\baselineskip}

@@

  \caption{Example Product Graph IR construction.}
  \label{fig:example-compilation}
\end{figure*}

@@

% Formally define automata
The RIR policy is of the form $r_1 \gg \dots \gg r_j$. While paths talk about the direction traffic flows through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated (i.e., route announcements flowing in the opposite direction). To capture this idea, for each regular expression $r_i$, we construct a deterministic finite state machine on the reversed regular expression. Each automaton is a tuple: ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all internal and external ASes, \ratul{ASes or locations?} $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon 2^{V \times V}$.
%
% Formally define product graph
The combined PGIR is a tuple: ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon 2^{V' \times V'}$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a vertex $v' = (v, \dots) \in V'$, we say $loc(v') = v$. We say a node $x$ shadows node $y$ in the product graph when $x \in V'$ and $y \in V'$ and $loc(x) = loc(y)$ but $x \neq y$. That is, $x$ shadows $y$ when they are different nodes in the PGIR that represent the same topology location.

@@

% Formal product construction
Let $a_i$ and $b_i$ denote states in the regular policy automata.
The PGIR is constructed by adding an edge from $v_1 = (x, a_1, \dots, a_m)$ to $v_2 = (y, b_1, \dots, b_m)$ whenever $\sigma_i(a_i, y) = b_i$ for each $i$ and $(x,y) \in E$ is a valid topology link.
%
Additionally, we add edges from the start node $s$ to any $v = (x, a_1, \dots, a_m)$ when $\sigma_i(q_{0_i}, x) = a_i$ for each $i$.
%
The preference function $P$ for node $v = (x, a_1, \dots, a_m)$ is defined as $P(v) = \set{i~\vert~a_i \in F_i} $. That is, it records which preferences are achieved for the policy automata for each node in the product construction.
%
Finally, there is an edge from each node in the PGIR such that $P(v) \neq \emptyset$ to the special end node $e$.

@@

\para{Minimization}
\ratul{This minimization section can probably go entirely, shrunken, or moved to the implementation}
Before checking if policies captured by the PGIR are safe, we minimize the PGIR to better scale our failure safety checks. The minimization is based on the observation that although every path in the PGIR is a valid path in the topology, we do not want to consider loops when configuring the network. In particular, BGP's loop prevention mechanism, by which an AS rejects any route for which it is already in the AS path, means that loops should never occur.\footnote{ Misconfigured aggregation can still lead to loops in BGP.}
%
However, the PGIR may contain many paths that form loops in the topology. For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path that BGP announcements might take, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
Since we are primarily concerned with guaranteeing user policy correctness and safety with respect to network failures, it is useful to eliminate cases that are not possible due to loops. For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
%
Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators~\ref{bib:todo}. For minimization, we compute the dominator set for each node in both the product graph $G$ with respect to the start node, and in the reversed product graph $G^R$ with respect to the end node. The following rules enable efficient minimization of the product graph.
%
\begin{itemize}
  \item Remove nodes that never reach the end node
  \item Remove nodes not reachable from the $start$ node
  \item Remove any node $x$ such that $x$ is dominated by $y$ in $G$ or $G^R$, and $loc(x) = loc(y)$
  \item Remove any edge from $x$ to $y$ in $G$ or $G^R$ if there is some node $z$ dominated by $y$ such that $loc(x) = loc(z)$
\end{itemize}
%
Repeated application of the above rules leads to removing the colored nodes and dashed lines in Figure~\ref{fig:example-compilation}.

@@

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$ and prefer them equally. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. In practice, we model the unknown AS topology as a special node in the product graph that generates a filter to match any sequence of ASes.

Finally, in our example, the external AS $W$ should prefer our internal router $A$ over $B$. In general, it is not possible to reliably control traffic entering the network beyond certain special cases. In this example, however, assuming our network and$W$ have an agreement to honor MEDs, the BGP MED attribute can influence $W$ to prefer $A$ over $B$. Additionally, the compiler can use the BGP no-export community to ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).

Figure~\ref{fig:config} shows the full configuration from the compilation example. To make configurations more compact and readable, the \sysname compiler rewrites configurations in ABGP, e.g., removing community tagging and matching when there is no ambiguity. Figure~\ref{config-min} shows minimized router configurations. \todo{figure was removed?}

@@

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix going through a router where an aggregate is advertised to go over a single link, even if there are several available in the toplogy. If that link fails, a black hole might be introduced. Fortunately, the PGIR captures this information precisely.
%, describing the how information meeting the policy flows over the topology.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty is that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures:
\begin{enumerate}
	\item Pick a random start-to-end path in PGIR.
	\item Remove all edges in the PGIR for the set of topological edges chosen.
	\item Repeat until no path exists.
\end{enumerate}
The number of PGIR paths that we are able to remove is a lower bound on the number of failures. Since each PGIR path results in a new edge-disjoint path through the topology (because of Step 2), and since the number of edge-disjoint paths through the topology is no more than the min-cut. \todo{clarify this}

>>>>

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}

  \hdr{\large Topology}{}
  \vspace*{-2\baselineskip}

@@

  \caption{Example Product Graph IR construction for policy: $(W \cdot A \cdot C \cdot D \cdot out) \gg (W \cdot B \cdot in^* \cdot out)$.}
  \label{fig:example-compilation}
\end{figure*}

@@

% Formally define automata
The RIR policy is an ordered sequence of regular expressions $r_1 \gg \dots \gg r_j$. While paths talk about the direction traffic flows through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated (i.e., route announcements flowing in the opposite direction). To capture this idea, for each regular expression $r_i$, we construct a deterministic finite state machine on the reversed regular expression. Each automaton is a tuple: ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all all abstraction locations (routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon 2^{V \times V}$.
%
% Formally define product graph
The combined PGIR is a tuple: ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon 2^{V' \times V'}$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a vertex $v' = (v, \dots) \in V'$, we say $loc(v') = v$. 
A node $x$ shadows node $y$ in the product graph when $x \in V'$ and $y \in V'$ and $loc(x) = loc(y)$ but $x \neq y$. That is, $x$ shadows $y$ when they are different nodes in the PGIR that represent the same topology location.

@@

% Formal product construction
Let $a_i$ and $b_i$ denote states in the regular policy automata.
The PGIR is constructed by adding an edge from $v_1 = (x, a_1, \dots, a_m)$ to $v_2 = (y, b_1, \dots, b_m)$ whenever $\sigma_i(a_i, y) = b_i$ for each $i$ and $(x,y) \in E$ is a valid topology link.
%
Additionally, we add edges from the start node $s$ to any $v = (x, a_1, \dots, a_m)$ when $\sigma_i(q_{0_i}, x) = a_i$ for each $i$.
%
The preference function $P$ for node $v = (x, a_1, \dots, a_m)$ is defined as $P(v) = \set{i~\vert~a_i \in F_i} $. That is, it records which preferences are achieved from each policy automaton for every node in the product construction.
%
Finally, there is an edge from each node in the PGIR such that $P(v) \neq \emptyset$ to the special end node $e$.

@@

\para{Minimization}
\ratul{This minimization section can probably go entirely, shrunken, or moved to the implementation}
Before checking if policies captured by the PGIR are safe, we minimize the PGIR to better scale our failure safety checks. The minimization is based on the observation that although every path in the PGIR is a valid path in the topology, we do not want to consider loops when configuring the network. In particular, BGP's loop prevention mechanism, by which an AS rejects any route for which it is already in the AS path, means that loops should never occur.\footnote{ Misconfigured aggregation can still lead to loops in BGP.}
%
However, the PGIR may contain many paths that form loops in the topology. For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path that BGP announcements might take, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
Since we are primarily concerned with guaranteeing user policy correctness and safety with respect to network failures, it is useful to eliminate cases that are not possible due to loops. 
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.
We use graph dominators~\ref{bib:todo} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any simple path between the start and end nodes.
A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators in practice~\ref{bib:todo}.
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node. 
For example, node $(1,1) W$ in Figure~\ref{fig:example-compilation} is never on a simple path to the end node since it most go through another $W$ in either case.
Colored nodes and dashed edges are removed after minimization since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

%A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators~\ref{bib:todo}. For minimization, we compute the dominator set for each node in both the product graph $G$ with respect to the start node, and in the reversed product graph $G^R$ with respect to the end node. The following rules enable efficient minimization of the product graph.
%
%\begin{itemize}
%  \item Remove nodes that never reach the end node
%  \item Remove nodes not reachable from the $start$ node
%  \item Remove any node $x$ such that $x$ is dominated by $y$ in $G$ or $G^R$, and $loc(x) = loc(y)$
%  \item Remove any edge from $x$ to $y$ in $G$ or $G^R$ if there is some node $z$ dominated by $y$ such that $loc(x) = loc(z)$
%\end{itemize}
%
%Repeated application of the above rules leads to removing the colored nodes and dashed lines in Figure~\ref{fig:example-compilation}.

@@

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. Any unknown AS topology is modelled as a special node in the product graph that generates a filter to match any sequence of ASes.

Finally, in our example, the external AS $W$ should prefer our internal router $A$ over $B$. In general, it is not possible to reliably control traffic entering the network beyond certain special cases. In this example, however, assuming our network and $W$ have an agreement to honor MEDs, the BGP MED attribute can influence $W$ to prefer $A$ over $B$. Additionally, the compiler can use the BGP no-export community to ensure that no other AS beyond $W$ can send us traffic. The compiler can perform a simple analysis to determine when it can utilize various BGP special attributes to ensure traffic enters the network in a particular way by looking at links in the product graph that cross from the internal topology to the external topology. 
%Aggregation is another tool commonly used to influence how traffic can enter the network, by relying on BGP's longest prefix matching default. Since aggregates are provided as additional user constraints, the compiler can also make use of this information to check that incoming preferences can be met given the aggregates specified (e.g., to ensure that all external ASes prefer to enter from one location over another).
Figure~\ref{fig:config} shows the full configuration from the compilation example.

To make configurations more compact and readable, the \sysname compiler makes several passes over the ABGP policy, removing community tagging and matching when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. After removing the communities, the two rules at router $B$ are merged into a single rule.

@@

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing of traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix to go over a single link before being aggregated, even if there are several available in the toplogy. If that link fails, a black hole might be introduced. Fortunately, the PGIR captures the information about both the policy and topology precisely.
%, describing the how information meeting the policy flows over the topology.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty is that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures: $i)$ Pick a random start-to-end path in PGIR, $ii)$ remove all edges in the PGIR for the set of topological edges chosen, and $iii)$ repeat until no such path exists.
The number of PGIR paths that we are able to remove is a lower bound on the number of failures required to disconnect the prefix from its aggregate. 
%Since each PGIR path results in a new edge-disjoint path through the topology (because of Step 2), and since the number of edge-disjoint paths through the topology is no more than the min-cut. \todo{clarify this}
